Report for Programming Assignment 2

### 2017-11-10 Programming-Assignment-2/PA_2-report.tex
PA-2 Update 6

\documentclass{article}
\usepackage[margin=0.73in]{geometry}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{float}
\usepackage{hyperref}

@@

\section*{Part b}
\subsection*{Question 1}
\begin{flushleft}
Below is the table representing the objective value of the optimal solution. The program used \texttt{CVXPY}, a convex programming module in Python.
\begin{center}
\begin{tabular}{|c|c|}
\hline
Penalty Parameter combination \((C_{1}, C_{2})\) & Objective Value \\
\hline
\hline
\((1, 1)\) & \(9.60436\)\\
\hline
\((1, 10)\) & \(30.26522\)\\
\hline
\((10, 1)\) & \(11.89106\)\\
\hline
\end{tabular}
\end{center}
\end{flushleft}

@@

>>>>

\documentclass{article}
\usepackage[margin=0.73in]{geometry}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{float}
\usepackage{bookmark}
\usepackage{hyperref}
\newcommand{\xtilde}{\widetilde{X}}
\newcommand{\indi}{\mathbb{I}}
\newcommand{\setS}{\mathcal{S}}
\newcommand{\real}{\mathbb{R}}
\newcommand{\lagL}{\mathcal{L}}
\newcommand{\minimize}{\mathop{\mathrm{minimize}}}
\newcommand{\minimizewrt}[1]{\underset{#1}{\minimize}}

@@

\section*{Instructions for the code}
\begin{itemize}
\item There are two files: \texttt{primal.py} and \texttt{dual.py}. \texttt{primal.py} concerns questions b.1, b.2 and b.4. \texttt{dual.py} concerns question b.3.
\item The programs used \texttt{CVXPY}, a convex programming module in Python.
\item To run these programs, enter \texttt{python3 <file>.py --C1 <C1-value> --C2 <C2-value>}. \texttt{<file>} can be any one of \texttt{primal} and \texttt{dual}.
\end{itemize}

\section*{Part a}
\subsection*{Question 1}
\begin{flushleft}
Strong duality holds for the below problem:
\begin{multline}
\minimizewrt{\beta \in \real^{p}, \hspace{0.5mm} \beta_{0} \in \real, \hspace{0.5mm} \xi \in \real^{n}} \frac{1}{2}||\beta||_{2}^{2} + C_{1}\sum_{i \in \setS_{1}} \xi_{i} + C_{2}\sum_{i \in \setS_{2}} \xi_{i} \\
\text{subject to } \xi_{i} \geq 0, \hspace{2mm} y_{i}(x_{i}^{T}\beta + \beta_{0}) \geq 1 - \xi_{i}, \hspace{2mm} \forall i \in \{1, 2, \ldots, n\}
\label{primal}
\end{multline}

The reasons are:
\begin{itemize}
\item The objective function is convex
\item The constraints are affine in the variables to be optimized over
\item Using the above two points and Slater's condition, we can tell \ref{primal} has strong duality.
\end{itemize}
\end{flushleft}

\subsection*{Question 2}
\begin{flushleft}
Define the Lagrangian \(\mathcal{L}\) as:
\begin{equation}
\label{lagrangian}
\displaystyle \lagL(\beta, \beta_{0}, \xi, \alpha, \mu) = \frac{1}{2}||\beta||_{2}^{2} + C_{1}\sum_{i \in \setS_{1}}\xi_{i} + C_{2}\sum_{i \in \setS_{2}}\xi_{i} + \sum_{i=1}^{n}\alpha_{i}(1 - \xi_{i} - y_{i}(x_{i}^{T}\beta + \beta_{0})) - \sum_{i=1}^{n}\mu_{i}(\xi_{i})
\end{equation}

By property of stationarity in \ref{lagrangian}, assuming \(\beta, \beta_{0}, \xi\) are solutions to the primal and \(\alpha, \mu\) are the solutions to the dual and using the stationarity sub-condition of the KKT conditions, we get:
\begin{gather}
\label{stationarity-beta}
\displaystyle \nabla_{\beta} \lagL = \beta + \sum_{i=1}^{n}\alpha_{i}(-y_{i}x_{i}) = 0 \implies \beta = \sum_{i=1}^{n}\alpha_{i}y_{i}x_{i} \\
\label{stationarity-beta-0}
\displaystyle \frac{\partial \lagL}{\partial \beta_{0}} = -\sum_{i}^{n}\alpha_{i}y_{i} = 0 \implies \sum_{i}^{n}\alpha_{i}y_{i} = 0 \\
\label{stationarity-xi}
\displaystyle \frac{\partial \lagL}{\partial \xi_{i}} = C_{1}\indi_{i \in \setS_{1}} + C_{2}\indi_{i \in \setS_{2}} - \alpha_{i} - \mu_{i} = 0 \implies \alpha_{i} = C_{1}\indi_{i \in \setS_{1}} + C_{2}\indi_{i \in \setS_{2}} - \mu_{i}
\end{gather}

Now using the complementary slackness sub-condition of the KKT conditions, we get:
\begin{gather}
\label{compl-slack-alpha}
\displaystyle \alpha_{i}(1 - \xi_{i} + y_{i}(x_{i}^{T}\beta + \beta_{0})) = 0 \hspace{4mm} \forall i \in \{1, 2, \ldots, n\} \\
\label{compl-slack-mu}
\displaystyle \mu_{i}\xi_{i} = 0 \hspace{4mm} \forall i \in \{1, 2, \ldots, n\}
\end{gather}

There is nothing specifically required to be written for the primal feasibility sub-condition, since the solution to the primal has to be naturally feasible. But using dual feasibility:
\begin{equation}
\label{dual-fease}
\alpha_{i}, \mu_{i} \geq 0 \hspace{4mm} i \in \{1, 2, \ldots, n\}
\end{equation}
\end{flushleft}

\subsection*{Question 3}
\begin{flushleft}
Note that \ref{stationarity-beta-0} can be written as \(y^{T}\alpha = 0\) by definition of the inner product. Similarly, construct a matrix:
\begin{equation}
\label{y-diag}
\mathop{diag}(y) = \begin{bmatrix} y_{1} & 0 & \ldots & 0 \\ 0 & y_{2} & & \vdots \\ \vdots & & \ddots & 0 \\ 0 & \ldots & 0 & y_{n} \end{bmatrix} 
\end{equation}

Using \ref{y-diag}, we can note that:
\begin{equation}
\label{x-tilde}
\xtilde = \mathop{diag}(y)X = \begin{bmatrix} y_{1}x_{1} \\ \vdots \\ y_{n}x_{n} \end{bmatrix}
\end{equation}

From \ref{stationarity-beta}, we can rewrite \(\beta\) as:
\begin{equation}
\label{beta-x-tilde}
\beta = \sum_{i=1}^{n} \alpha_{i}\xtilde_{i} = \xtilde^{T}\alpha \implies ||\beta||_{2}^{2} = \beta^{T}\beta = \alpha^{T}\xtilde\xtilde^{T}\alpha
\end{equation}

Using \ref{dual-fease} and \ref{stationarity-xi}, we can tell that:
\begin{equation}
\label{alpha-S}
0 \leq \alpha_{i} \leq C_{1}\indi_{i \in \setS_{1}} + C_{2}\indi_{i \in \setS_{2}} \implies 0 \leq \alpha_{\setS_{j}} \leq C_{j}\mathbf{1} \hspace{2mm} \forall j \in \{1, 2\}
\end{equation}

The above step in \ref{alpha-S} uses the fact that if \(i \in \setS_{1}\), then \(i \notin \setS_{2}\). \(\alpha_{\setS_{j}}\) denotes the vector of \(\alpha_{i}\)s such that \(i \in \setS_{j}\).

\end{flushleft}

\section*{Part b}
\subsection*{Question 1}
\begin{flushleft}
Below is the table representing the objective value of the optimal solution.
\begin{center}
\begin{tabular}{|c|c|}
\hline
Penalty Parameter combination \((C_{1}, C_{2})\) & Objective Value \\
\hline
\hline
\((1, 1)\) & \(9.60436\)\\
\hline
\((1, 10)\) & \(30.26522\)\\
\hline
\((10, 1)\) & \(11.89106\)\\
\hline
\end{tabular}
\end{center}
\end{flushleft}

@@

\subsection*{Question 4}
The total classification error calculated in a zero-one loss manner is below:
\begin{center}
\begin{tabular}{|c|c|}
\hline
Penalty Parameter combination \((C_{1}, C_{2})\) & Weighted Classification Error \\
\hline
\hline
\((1, 1)\) & \(3\) \\
\hline
\((1, 10)\) & \(16\) \\
\hline
\((10, 1)\) & \(4\) \\
\hline
\end{tabular}
\end{center}
